---
title: "Lesson Template Notebook, Completed"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output:
  github_document: default
  html_notebook: default
---

## Introduction
This is a sample notebook that illustrates the notebook template that should be used for SLU DSS lessons.

## Load Dependencies
The following code loads the package dependencies for our analysis from the `tidyverse`:

```{r package-load-1}
# tidyverse packages
library(dplyr)    # data wrangling
library(ggplot2)  # data plotting
library(readr)    # work with csv files
```

Now you try loading two additional packages - `here` (for working with file paths) and `skimr` (for generating descriptive statistics):

```{r package-load-2}
# other packages
library(here)    # work with file paths
library(skimr)   # descriptive statistics
```

## Load Data
We'll use a data set describing asthma rates in St. Louis:

```{r load-asthma-data}
asthma <- read_csv(here("data", "stlAsthma.csv"))
```

We also want to use a data set describing un-insurance rates in St. Louis (i.e. the number of people without health insurance). Practice loading data using the `read_csv()` and `here()` functions. Load your data into an object called `unins`:

```{r load-unins-data}
unins <- read_csv(here("data", "stlInsurance.csv"))
```

## Explore Data
### Listing Variables
We'll start by using the `glimpse()` function from `dplyr` to list all of the variables and their formats:

```{r list-asthma-vars}
glimpse(asthma)
```

The key variables of interest here are the three variables that begin with `pct` - they include an estimated asthma prevalence for each census tract (`pctAsthma`) as well as high and low values based on a confidence interval for the main estimate. Each of these three variables is stored as a `dbl` (or "double") variable, meaning they contain numeric data suitable for statistical analysis and plotting.

Use the `glimpse()` function to create a similar list for the un-insurance data:

```{r list-unins-vars}
glimpse(unins)
```

### Descriptive Statistics
The `skimr` package provides an excellent descriptive statistics function called `skim()` that we can use to get quick numeric and visual(!) summaries of our data:

```{r descriptive-stats-asthma}
skim(asthma)
```

`skim()` segregates variables based on their type, unlike the base `R` `summary()` function. It also provides a small histogram for each distribution, which allows for easy comparison, and other summary statistics both for individual variables and the data set as a whole.

Use the `skim()` function to produce descriptive statistics for the un-insurance data:

```{r descriptive-stats-unins}
skim(unins)
```

### Histogram
Finally, we'll create a histogram of the main variable of interest - `pctAsthma`. We'll use `ggplot2` to build the plot:

```{r asthma-histogram}
ggplot() +
  geom_histogram(data = asthma, mapping = aes(pctAsthma))
```

We can see the a large number of census tracts have low asthma rates (below 10%). However, there are a sizeable cluster of tracts with rates between 12 and 15% with several tracts that have even higher rates.

Now it is your turn again! Create a histgoram of the un-insurance rate estimate for St. Louis:

```{r unins-histogram}
ggplot() +
  geom_histogram(data = unins, mapping = aes(pctUnins))
```
