---
title: "Lesson Template Challenge, Completed"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output:
  github_document: default
  html_notebook: default
---

## Introduction
This is a sample notebook that illustrates the notebook template that should be used for SLU DSS lesson challenges. We've provided a little bit of structure, including a short example of the literate programming style, to help get you started!

## Tasks

1. Load the tabular data set `stlSmoking.csv` into your enviornment
2. Make a list of the included variables using `dplyr`
3. Calculate descriptive statistics for the included variables using `skmir`
4. Create a histogram using `ggplot2`

## Load Dependencies
The following code loads the package dependencies.

```{r package-load}
# tidyverse packages
library(dplyr)    # data wrangling
library(ggplot2)  # data plotting
library(readr)    # work with csv files

# other packages
library(here)    # work with file paths
library(skimr)   # descriptive statistics
```

## Load Data

We'll use a data set describing smoking rates in St. Louis:

```{r load-data}
smoking <- read_csv(here("data", "stlSmoking.csv"))
```

## Explore Data
### Listing Variables
We'll start by using the `glimpse()` function from `dplyr` to list all of the variables and their formats:

```{r list-vars}
glimpse(smoking)
```

The key variables of interest here are the three variables that begin with `pct` - they include an estimated smoking prevalence for each census tract (`pctSmoke`) as well as high and low values based on a confidence interval for the main estimate. Each of these three variables is stored as a `dbl` (or "double") variable, meaning they contain numeric data suitable for statistical analysis and plotting.

### Descriptive Statistics
Next, we'll use the `skimr` package to generate descriptive statistics and visual summaries of our data:

```{r descriptive-stats}
skim(smoking)
```

For the main estimate, the average smoking prevalence is 26.40% with a range of 14.30% to 37.9%, meaning that about a quarter of the adult population of typical census tract smokes with a maximum of more than a third. The true population value of the estimate is likely to be between 23.03% and 29.86%.

### Histogram
Finally, we'll create a histogram of the main variable of interest - `pctSmoke`. We'll use `ggplot2` to build the plot:

```{r histogram}
ggplot() +
  geom_histogram(data = smoking, mapping = aes(pctSmoke))
```

We can see that the distribution of smoking prevalence is bimodal - there are peaks at around 20% of the adult population and again at around 28% or 29% of the adult population.
